Evaluating Deep Learning Methods for Tokenization of Space-less texts in Old French

### 2019-06-05 article/article.tex
Updated article

Tokenization of space-less strings is a task that is specifically difficult for computer when compared to "whathumancando". \textit{Scripta continua} is a writing phenomenon where words are be separated by spaces and has disappeared around the 8th century (see \citet{zanna1998lecture}). Never the less, spacing can be somewhat erratic in later centuries writings, as show by Figure \ref{fig:4lines}, a document from the 13th century. In the context of text mining of HTR or OCR output, lemmatization and tokenization of medieval western languages can be a pre-processing step for further research to sustain analyses such as authorship attribution \textbf{CITE JBCAMPS ?}.

@@

While all model using CNN shows improvement over the baseline, the model definitely does not outperform it by a huge margin (\textless 0.02 FScore). And for a reason : the baseline already performs nearly perfectly on the test corpus. The dictionary attack using n-grams did actually perform well. As a result, we wanted to compare how both models would perform on a secondary test corpus composed by texts that were not used in training: indeed, the training, dev and test corpus share the same texts while not sharing the same inputs. As a result, we created a new corpus with 4 texts and 742 samples : the diplomatic edition of the Graal \citet{graal}, a Passion and a Vie de Saint Leger \citet{old_french_corpus}, a Vie de Saint Thibaut \citet{theobaldus}. No noise characters and no random keeping of words were applied.

@@

\begin{table}[!ht]
\centering
\begin{tabular}{lllllll}
\hline
 & Accuracy & Precision & Recall & FScore & WB as WC & WC as WB \\ \hline
Baseline & 0.882 & 0.893 & 0.808 & 0.838 & 3658 & 644 \\
CNN P & \textbf{0.957} & \textbf{0.948} & \textbf{0.944} & \textbf{0.945} & \textbf{854} & \textbf{723} \\ \hline
\end{tabular}
\caption{Scores over the unknown dataset. WB as WC and WC as WB are taken from the confusion matrices}
\label{tab:scores_unknown}
\end{table}

@@

@@

The medieval Latin corpora is based on the project Formulae - Litterae - Chartae's open data \citet{formulae} for its training, evaluation and test sets; the unknown corpora is based on three texts from the Monumenta Germanica that are from early to late medieval period (Andreas Agnellus, Manegaldus, Theodoricus de Niem) and are drawn from the Corpus Corporum Project. Both corpus were generated without noise and word keeping, with a maximum sample size of 150 characters. The data presents some MUFI characters but still look like mostly normalized editions, unlike the Old French data.

@@

>>>>

Tokenization of space-less strings is a task that is specifically difficult for computer when compared to "whathumancando". \textit{Scripta continua} is a writing phenomenon where words are be separated by spaces and has disappeared around the 8th century (see \citet{zanna1998lecture}). Never the less, spacing can be somewhat erratic in later centuries writings as \citet{stutzmann} has shown (\textit{cf.} Figure \ref{fig:4lines}, a document from the 13th century, displays this kind of situation). In the context of text mining of HTR or OCR output, lemmatization and tokenization of medieval western languages can be a pre-processing step for further research to sustain analyses such as authorship attribution or simply allow full-text search.

@@

While all model using CNN shows improvement over the baseline, the model definitely does not outperform it by a huge margin (\textless 0.02 FScore). And for a reason : the baseline already performs nearly perfectly on the test corpus. The dictionary attack using n-grams did actually perform well. As a result, we wanted to compare how both models would perform on a secondary test corpus composed by texts that were not used in training: indeed, the training, dev and test corpus share the same texts while not sharing the same inputs. As a result, we created a new corpus with 4 texts and 742 samples : the diplomatic edition of the \textit{Graal} (\citet{graal}), a \textit{Passion} and a \textit{Vie de Saint Leger}  (\citet{old_french_corpus}), a \textit{Vie de Saint Thibaut} (\citet{theobaldus}). No noise characters and no random keeping of words were applied.

@@

\begin{table}[!ht]
\centering
\begin{tabular}{lllllll}
\hline
 & Accuracy & Precision & Recall & FScore & WB FN & WB FP \\ \hline
Baseline & 0.882 & 0.893 & 0.808 & 0.838 & 3658 & 644 \\
CNN P & \textbf{0.957} & \textbf{0.948} & \textbf{0.944} & \textbf{0.945} & \textbf{854} & \textbf{723} \\ \hline
\end{tabular}
\caption{Scores over the unknown dataset. FN = False Negative, FP = False Positive}
\label{tab:scores_unknown}
\end{table}

@@

\textbf{Example:}

\begin{itemize}
    \item Input : operecuperemdeberemqueprofecto
    \item Output : opere cuperem deberemque profecto

\end{itemize}

@@

The medieval Latin corpora is based on the project Formulae - Litterae - Chartae's open data (\citet{formulae}) for its training, evaluation and test sets; the unknown corpora is based on three texts from the Monumenta Germanica (\citet{germanica}) that are from early to late medieval period (Andreas Agnellus, Manegaldus, Theodoricus de Niem) and are drawn from the Corpus Corporum Project. Both corpus were generated without noise and word keeping, with a maximum sample size of 150 characters. The data presents some MUFI characters but still look like mostly normalized editions, unlike the Old French data.

@@

\textbf{Example:}

\begin{itemize}
    \item Input : nonparvamremtibi
    \item Output : non parvam rem tibi

\end{itemize}

\section{Annex 4: Scores on Latin epigraphic corpora}

\begin{table}[H]
\centering
\begin{tabular}{llllllll}
\hline
 & Corpus & Accuracy & Precision & Recall & FScore & WB FN & WB FP \\ \hline
Baseline & Test & 0.956 & 0.935 & 0.943 & 0.939 & 2646 & 3547 \\
CNN P & Test & \textbf{0.987} & \textbf{0.983} & \textbf{0.979} & \textbf{0.981} & \textbf{1149} & \textbf{722} \\ \hline
Baseline & Test Uppercase & 0.956 & 0.935 & 0.942 & 0.938 & 2664 & 3457 \\
CNN P & Test Uppercase & \textbf{0.979} & \textbf{0.972} & \textbf{0.967} & \textbf{0.969} & \textbf{1715} & \textbf{1275} \\ \hline
Baseline & Unknown & 0.879 & 0.834 & 0.817 & 0.825 & 8693 & 11332 \\
CNN P & Unknown & 0.953 & 0.939 & 0.926 & 0.932 & 4689 & 3112 \\ \hline
Baseline & Unknown Uppercase & 0.879 & 0.834 & 0.817 & 0.825 & 8693 & 11332 \\
CNN P & Unknown Uppercase & 0.936 & 0.914 & 0.902 & 0.908 & 6152 & 4464 \\ \hline
\end{tabular}
\caption{Scores over the Latin epigraphic datasets. FN = False Negative, FP = False Positive}
\label{tab:epigraphyc_latin}
\end{table}

The epigraphic Latin corpora is based on the Epigraphic Database Heidelberg open data \citet{formulae} for its training, evaluation and test sets (HD000001-HD010000 and HD010001-HD020000 from \citet{edh}) while the corpus of unknown is drawn from an automatic conversion of the Pompei Inscriptions (\citet{pompei}). We also decided to evaluate both the baseline and the model on uppercased data, as it would normally be the state the text would be found in. Each of the corpora presents a high number of unresolved abbreviations (\textit{ie.} one letter words). Both corpus were generated without noise and word keeping, with a maximum sample size of 150 characters. The data presents some polytonic Greek characters, some sample being only in Greek.

While CNN is influenced by
\textbf{Statistics:}

\begin{itemize}
\item Number of training examples: 46,423
\item Number of evaluation examples: 5,802
\item Number of testing examples: 5,804
\item Number of classes in testing examples: 107,963 WC; 31,900 WB
\item Number of classes in unknown examples: 127,268 WC; 38,055 WB
\end{itemize}

\textbf{Example:}

\begin{itemize}
    \item Input : DnFlClIuliani
    \item Output : D n Fl Cl Iuliani
\end{itemize}
